=enter title here=

### 2021-10-09 paper/agujournaltemplate.tex
Fix noise question

The DWT and MODWT methods must be used on a continuous time series, without gaps in the recordings. To deal with the gaps in the GNSS recordings, we simply replace the missing values by Gaussian noise with mean zero and standard deviation equal to the standard deviation of the whole time series. We verify how the wavelet details may be affected by looking at a GPS time series without missing values and comparing the wavelet details with and without removing some data points. Station PGC5 has recorded during 1390 days between 2009 and 2013, without any missing values. We first computed the wavelet details without missing values. Then, we removed ten neighboring missing values, replaced them by Gaussian noise, and computed the wavelet details with the replaced values. Figure 2 shows a comparison of the two wavelet details for two different locations of the missing values. We can see that there are visible differences in the time series itself, and in the details at the smallest levels of the wavelet decomposition. However, the differences between the wavelet details with and without missing values get smaller and smaller with increasing levels the details, and are barely visible for the levels we are mostly interested in (levels 6 to 8). We thus conclude that we can easily replace the missing values in the GNSS time series without introducing false detections of slow slip events. \\

\begin{figure}
\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
\caption{Bottom: Data from GPS station PGC5 without missing values (black) and with missing values replaced by Gaussian noise (red) for two locations of the missing values (left and right). Bottom to top: Corresponding eight details and smooths of the wavelet composition for the original data (black) and for the missing values replaced by Gaussian noise (red).}
\label{pngfiguresample}
\end{figure}

>>>>

The DWT and MODWT methods must be used on a continuous time series, without gaps in the recordings. To deal with the gaps in the GNSS recordings, we simply replace the missing values by the sum of a straight line and a Gaussian noise component with mean zero and standard deviation equal to the standard deviation of the whole time series. The straight line starts at the mean of the five days before the gap and ends at the mean of the five days after the gap. We verify how the wavelet details may be affected by looking at a GPS time series without missing values and comparing the wavelet details with and without removing some data points. Station PGC5 has recorded during 1390 days between 2009 and 2013, without any missing values. We first computed the wavelet details without missing values. Then, we removed ten neighboring missing values, replaced them by the sum of the straight line and the Gaussian noise component, and computed the wavelet details with the replaced values. Figure 2 shows a comparison of the two wavelet details for two different locations of the missing values. We can see that there are visible differences in the time series itself, and in the details at the smallest levels of the wavelet decomposition. However, the differences between the wavelet details with and without missing values get smaller and smaller with increasing levels the details, and are barely visible for the levels we are mostly interested in (levels 6 and above). We thus conclude that we can easily replace the missing values in the GNSS time series without introducing false detections of slow slip events. \\

\begin{figure}
\noindent\includegraphics[width=5cm, trim={0cm 0cm 0cm 0cm},clip]{figures/DS_10.eps}
\caption{Bottom: Data from GPS station PGC5 without missing values (black) and with missing values replaced by the sum of a straight line and a Gaussian noise component (red) for two locations of the missing values (left and right). Bottom to top: Corresponding ten details and smooths of the wavelet composition for the original data (black) and for the missing values replaced by the sum of a straight line and a Gaussian noise component (red).}
\label{pngfiguresample}
\end{figure}
